Graph Storage: How good is CSR really?

### 2017-12-10 Seminar Paper/paper.tex
Presentation is added, some additions on the seminar paper

\section{Evaluation}
All three approches are tested on different datasets which are generated randomly. The number of users is 1,000,000. To make it as a real social network, the number of neighbors differs from 50 to 200. The neighbors are actually selected randomly. In the following subsections, I discuss in which scenario CSR gives better performance in terms of latency and memory consumption. My test machine is a single socket server computer equipped with Intel(R) Core(TM) i7-3930K CPU 6-core @ 3.20GHz and 64 GB of main memory. This server has 12 hardware cores (with hyperthreading enabled).

@@

 \subsubsection*{Simple Update}
   Assume that the directed graph has \textit{V} nodes and \textit{E} edges. To keep the structure of CSR, we need to allocate a new array with \textit{E}+\textit{1} edges for each newly added edge. Let's call the new array as \texttt{edges'} and the new edge is \textit{e=(v,u)}. First of all, I copy all the content of the array \texttt{edges} from \texttt{offsets[1]} to \texttt{offsets[v+1]} and the rest content of the array \texttt{edges} to the array \texttt{edges'} starting from \texttt{offsets[v+1] + 1} index. Then I set \texttt{edges'[offsets[v+1]]=u}. At the end, I increment all the elements from \texttt{offsets[v+1]} to \texttt{offsets[V+1]}. To copy the content of array, I use \texttt{std::memcpy} function which is better than \texttt{std::memmove} function. 
 \subsubsection*{Light Update}
 As it is shown from the results of \textbf{\textit{Simple Update}}, it is the most time consuming operation for CSR format. A new update structure -\textbf{\textit{Light Update}}- is a much faster update method compared to \textbf{\textit{Simple Update}}.\\
 The new approach is built on top of the previous approach. As an addition, the structure of CSR format should be modified in this method. Assume that we are given a directed graph which has \textit{V} nodes and \textit{E} edges. \texttt{edges} array should have more than |\textit{E}|, which I call it as \textit{additional space}. For each new upcoming edge, CSR doesn't need to allocate a new big \texttt{edges} array to store a new one. Let's call the new edge as \textit{e=(v,u)}. It does a new allocation whenever \texttt{edges} array hits the capacity. If it is already full, the update is exactly the same as in \textbf{\textit{Simple Update}}. If the \texttt{edges} array still has some space to store new edges, I copy all the array elements from \texttt{offsets[v+1]} into \texttt{edges} array starting from \texttt{offsets[v+1] + 1} index. As there is intersection between source and destination, I use \texttt{std::memmove} function to do this copy operation. Then I set the corresponding element to \texttt{u} and increment corresponding elements of \texttt{offsets} array as it is in \textbf{\textit{Simple Update}}. As it is shown, CSR doesn't do new allocation and copy all the array for each update. The speed-up comes along with having very few new allocations and copying.\\
\begin{center}
\includegraphics[scale = 0.5]{time_per_update}\\
\textbf{Figure 7}: Latencies of both updates for the graph with 1,000,000 nodes.
\end{center}

>>>>

\section{Evaluation}
All three approches are tested on different datasets which are generated randomly. The number of users is 1,000,000. To make it as a real social network, the number of neighbors differs from 50 to 200. The neighbors are actually selected randomly. The graph that I generated is directed and weighted, as I need to run Dijkstra Algorithm. In the following subsections, I discuss in which scenario CSR gives better performance in terms of latency and memory consumption. My test machine is a single socket server computer equipped with Intel(R) Core(TM) i7-3930K CPU 6-core @ 3.20GHz and 64 GB of main memory. This server has 12 hardware cores (with hyperthreading enabled).

@@

 \subsubsection*{Simple Update}
   Assume that the directed graph has \textit{V} nodes and \textit{E} edges. To keep the structure of CSR, we need to allocate a new array with \textit{E}+\textit{1} edges for each newly added edge. Let's call the new array as \texttt{edges'} and the new edge is \textit{e=(v,u)}. First of all, I copy all the content of the array \texttt{edges} from \texttt{offsets[1]} to \texttt{offsets[v+1]} and the rest content of the array \texttt{edges} to the array \texttt{edges'} starting from \texttt{offsets[v+1] + 1} index. Then I set \texttt{edges'[offsets[v+1]]=u}. At the end, I increment all the elements from \texttt{offsets[v+1]} to \texttt{offsets[V+1]}. To copy the content of array, I use \texttt{std::memcpy} function which is better than \texttt{std::memmove} function. 
 \subsubsection*{Light Update}
 As it is shown from the results of \textbf{\textit{Simple Update}}, it is the most time consuming operation for CSR format. A new update structure -\textbf{\textit{Light Update}}- is a much faster update method compared to \textbf{\textit{Simple Update}}.\\
 The new approach is built on top of the previous approach. As an addition, the structure of CSR format should be modified in this method. Assume that we are given a directed graph which has \textit{V} nodes and \textit{E} edges. \texttt{edges} array should have more than |\textit{E}|, which I call it as \textit{additional space}. For each new upcoming edge, CSR doesn't need to allocate a new big \texttt{edges} array to store a new one. Let's call the new edge as \textit{e=(v,u)}. It does a new allocation whenever \texttt{edges} array hits the capacity. If it is already full, the update is exactly the same as in \textbf{\textit{Simple Update}}. If the \texttt{edges} array still has some space to store new edges, I copy all the array elements from \texttt{offsets[v+1]} into \texttt{edges} array starting from \texttt{offsets[v+1] + 1} index. As there is intersection between source and destination, I use \texttt{std::memmove} function to do this copy operation. Then I set the corresponding element to \texttt{u} and increment corresponding elements of \texttt{offsets} array as it is in \textbf{\textit{Simple Update}}. As it is shown, CSR doesn't do new allocation and copy all the array for each update. The speed-up comes along with having very few new allocations and copying. The latency for \textbf{\textit{Light Update}} can simply be twice as less if the graph is unweighted, as I am also updating the weights array for each new edge.\\
\begin{center}
\includegraphics[scale = 0.5]{time_per_update}\\
\textbf{Figure 7}: Latencies of both updates for the graph with 1,000,000 nodes.
\end{center}
